---
title: "DECAF API Client R Library Tutorial"
output:
  html_document:
    toc: true
    toc_depth: 3
---

```{r, include = FALSE}
options(width=100)
```

## Creating DECAF Client

A DECAF client is an instance of `rdecaf::DecafClient` R6 class:

```{r, include = TRUE}
client <- rdecaf::DecafClient$new(
    url = "https://telosinvest.decafhub.com",
    credentials = list(username = username, password = password)
)
```

The `client` value has some members that can be used to query remote
DECAF API interfaces or get information about the client itself:

```{r}
names(client)
```

For example:

```{r}
client$url
```

... or:

```{r}
client$info()
```

## Bare Client

Using the `bare` client, we can place any kind of API request to any
DECAF API interface. The `bare` client is an instance of `HttpClient`
R6 class from the
[crul](https://cran.r-project.org/web/packages/crul/index.html)
library. It has all the necessary configuration made by the
`rdecaf::DecafClient` constructor. We strongly suggest you to check
the ["crul introduction" R
Vignette](https://cran.r-project.org/web/packages/crul/vignettes/crul.html)
and other accompanying `crul` documentation material to learn more
about it.

To demonstrate the common use pattern, we will request the remote
DECAF Barista API version information now:

```{r, include = TRUE}
version_response <- client$bare$get("/api/version/")
```

The `version_response` value is an instance of `HttpResponse` R6 class
of [crul](https://cran.r-project.org/web/packages/crul/index.html)
library:

```{r, include = TRUE}
class(version_response)
```

It has quite a few members:

```{r}
names(version_response)
```

In the following subsections, we will see some of these members which
we will end up using in our programs.

### Checking the Status Code

The status code for a successful HTTP request is between `200` and
`399`. We can check the success code of the response as follows:

```{r}
version_response$status_code
```

Also, we can check if the request was successful:

```{r}
version_response$success()
```

An undesired case is to attempt to get a non-existing HTTP resource:

```{r}
client$bare$get("/api/hebele/hubele/")$status_code
```

```{r}
client$bare$get("/api/hebele/hubele/")$success()
```

### Checking the Headers and Content Type

You can retrieve all response headers:

```{r}
version_response$response_headers
```

One of them is the content type that you may need:

```{r}
version_response$response_headers$"content-type"
```

### Getting the Response Content

`version_response` value carries the response content. But it is
provided in the raw format. This is normal as it can be of any content
type including binary data.

```{r}
version_response$content
```

`crul` can parse the response content:

```{r}
version_response$parse()
```

Note the encoding warning! `curl` did not know which encoding to
use. `utf-8` is the most common textual encoding DECAF is
using. Actually, anything other than `utf-8` would be a bug. So, you
can safely assume `utf-8` encoding when attempting to parse such
content:

```{r}
version_response$parse(encoding = "utf-8")
```

... which would give no encoding warnings.

You can then parse the content into a suitable R data structure:

```{r}
jsonlite::fromJSON(version_response$parse(encoding = "utf-8"))
```

### Summary

We can use the following one-liner to consume a remote DECAF API
endpoint that has a JSON content-type:

```{r}
jsonlite::fromJSON(client$bare$get("/api/version/")$parse(encoding="utf-8"))
```

However, this is a too optimistic. What if the request is
unsuccessful? Let's write a small function that will serve as a
helper:

```{r}
getJSON <- function(client, ...) {
  response <- client$bare$get(...)
  if (response$success()) {
	jsonlite::fromJSON(response$parse(encoding = "utf-8"))
  } else {
    stop(sprintf("Error while requesting %s. Status code: %s. Error: %s", response$url, response$status_code, response$parse(encoding = "utf-8")))
  }
}
```

Happy path:

```{r}
getJSON(client, "/api/version/")
```

Not so happy path:

```{r, error = TRUE}
getJSON(client, "/api/hebele/hubele/")
```

... but at least we have better error message that informs us for our
debugging.
